\chapter*{Bibliography}
\addcontentsline{toc}{chapter}{Bibliography}

\section*{Primary References}

\begin{enumerate}
\item Bloom, B. H. (1970). ``Space/time trade-offs in hash coding with allowable errors.'' \textit{Communications of the ACM}, 13(7), 422-426.

\item Broder, A., \& Mitzenmacher, M. (2004). ``Network applications of Bloom filters: A survey.'' \textit{Internet Mathematics}, 1(4), 485-509.

\item Mitzenmacher, M., \& Upfal, E. (2017). \textit{Probability and computing: Randomization and probabilistic techniques in algorithms and data analysis}. Cambridge University Press.

\item Fan, L., Cao, P., Almeida, J., \& Broder, A. Z. (2000). ``Summary cache: A scalable wide-area web cache sharing protocol.'' \textit{IEEE/ACM Transactions on Networking}, 8(3), 281-293.
\end{enumerate}

\section*{Cryptography and Hash Functions}

\begin{enumerate}
\setcounter{enumi}{4}
\item Katz, J., \& Lindell, Y. (2020). \textit{Introduction to modern cryptography} (3rd ed.). CRC Press.

\item Menezes, A. J., Van Oorschot, P. C., \& Vanstone, S. A. (1996). \textit{Handbook of applied cryptography}. CRC Press.

\item Rogaway, P., \& Shrimpton, T. (2004). ``Cryptographic hash-function basics: Definitions, implications, and separations for preimage resistance, second-preimage resistance, and collision resistance.'' In \textit{Fast Software Encryption} (pp. 371-388). Springer.

\item Aumasson, J. P. (2019). \textit{Serious cryptography: A practical introduction to modern encryption}. No Starch Press.

\item Bellare, M., \& Rogaway, P. (2006). ``The security of triple encryption and a framework for code-based game-playing proofs.'' In \textit{Advances in Cryptology - EUROCRYPT 2006} (pp. 409-426).
\end{enumerate}

\section*{Privacy and Searchable Encryption}

\begin{enumerate}
\setcounter{enumi}{9}
\item Song, D. X., Wagner, D., \& Perrig, A. (2000). ``Practical techniques for searches on encrypted data.'' In \textit{Proceedings of the 2000 IEEE Symposium on Security and Privacy} (pp. 44-55).

\item Curtmola, R., Garay, J., Kamara, S., \& Ostrovsky, R. (2006). ``Searchable symmetric encryption: Improved definitions and efficient constructions.'' In \textit{Proceedings of the 13th ACM Conference on Computer and Communications Security} (pp. 79-88).

\item Cash, D., Jarecki, S., Jutla, C., Krawczyk, H., Roşu, M. C., \& Steiner, M. (2013). ``Highly-scalable searchable symmetric encryption with support for Boolean queries.'' In \textit{Advances in Cryptology - CRYPTO 2013} (pp. 353-373).

\item Islam, M. S., Kuzu, M., \& Kantarcioglu, M. (2012). ``Access pattern disclosure on searchable encryption: Ramification, attack and mitigation.'' In \textit{Network and Distributed System Security Symposium} (Vol. 20, p. 12).

\item Cash, D., Grubbs, P., Perry, J., \& Ristenpart, T. (2015). ``Leakage-abuse attacks against searchable encryption.'' In \textit{Proceedings of the 22nd ACM SIGSAC Conference on Computer and Communications Security} (pp. 668-679).

\item Grubbs, P., McPherson, R., Naveed, M., Ristenpart, T., \& Shmatikov, V. (2016). ``Breaking web applications built on top of encrypted data.'' In \textit{Proceedings of the 2016 ACM SIGSAC Conference on Computer and Communications Security} (pp. 1353-1364).

\item Pappas, V., Krell, F., Vo, B., Kolesnikov, V., Malkin, T., Choi, S. G., ... \& Bellovin, S. (2014). ``Blind seer: A scalable private DBMS.'' In \textit{2014 IEEE Symposium on Security and Privacy} (pp. 359-374).

\item Demertzis, I., Papadopoulos, D., Papamanthou, C., \& Shintre, S. (2020). ``SEAL: Attack mitigation for encrypted databases via adjustable leakage.'' In \textit{29th USENIX Security Symposium} (pp. 2433-2450).
\end{enumerate}

\section*{Private Information Retrieval}

\begin{enumerate}
\setcounter{enumi}{17}
\item Chor, B., Goldreich, O., Kushilevitz, E., \& Sudan, M. (1995). ``Private information retrieval.'' In \textit{Proceedings of the 36th Annual Symposium on Foundations of Computer Science} (pp. 41-50).

\item Goldreich, O., \& Ostrovsky, R. (1996). ``Software protection and simulation on oblivious RAMs.'' \textit{Journal of the ACM}, 43(3), 431-473.

\item Stefanov, E., Van Dijk, M., Shi, E., Fletcher, C., Ren, L., Yu, X., \& Devadas, S. (2013). ``Path ORAM: An extremely simple oblivious RAM protocol.'' In \textit{Proceedings of the 2013 ACM SIGSAC Conference on Computer and Communications Security} (pp. 299-310).

\item Patel, S., Persiano, G., Raykova, M., \& Yeo, K. (2019). ``PanORAMa: Oblivious RAM with logarithmic overhead.'' In \textit{2018 IEEE 59th Annual Symposium on Foundations of Computer Science} (pp. 871-882).
\end{enumerate}

\section*{Homomorphic Encryption}

\begin{enumerate}
\setcounter{enumi}{21}
\item Gentry, C. (2009). ``A fully homomorphic encryption scheme.'' PhD thesis, Stanford University.

\item Brakerski, Z., Gentry, C., \& Vaikuntanathan, V. (2014). ``(Leveled) fully homomorphic encryption without bootstrapping.'' \textit{ACM Transactions on Computation Theory}, 6(3), 1-36.

\item Cheon, J. H., Kim, A., Kim, M., \& Song, Y. (2017). ``Homomorphic encryption for arithmetic of approximate numbers.'' In \textit{Advances in Cryptology - ASIACRYPT 2017} (pp. 409-437).
\end{enumerate}

\section*{Differential Privacy}

\begin{enumerate}
\setcounter{enumi}{24}
\item Dwork, C., McSherry, F., Nissim, K., \& Smith, A. (2006). ``Calibrating noise to sensitivity in private data analysis.'' In \textit{Theory of Cryptography Conference} (pp. 265-284).

\item Dwork, C., \& Roth, A. (2014). ``The algorithmic foundations of differential privacy.'' \textit{Foundations and Trends in Theoretical Computer Science}, 9(3-4), 211-407.

\item Vadhan, S. (2017). ``The complexity of differential privacy.'' In \textit{Tutorials on the Foundations of Cryptography} (pp. 347-450). Springer.

\item Abadi, M., Chu, A., Goodfellow, I., McMahan, H. B., Mironov, I., Talwar, K., \& Zhang, L. (2016). ``Deep learning with differential privacy.'' In \textit{Proceedings of the 2016 ACM SIGSAC Conference on Computer and Communications Security} (pp. 308-318).
\end{enumerate}

\section*{Information Theory}

\begin{enumerate}
\setcounter{enumi}{28}
\item Cover, T. M., \& Thomas, J. A. (2006). \textit{Elements of information theory} (2nd ed.). Wiley-Interscience.

\item MacKay, D. J. (2003). \textit{Information theory, inference and learning algorithms}. Cambridge University Press.

\item Shannon, C. E. (1948). ``A mathematical theory of communication.'' \textit{The Bell System Technical Journal}, 27(3), 379-423.
\end{enumerate}

\section*{Probabilistic Data Structures}

\begin{enumerate}
\setcounter{enumi}{31}
\item Cormode, G., \& Muthukrishnan, S. (2005). ``An improved data stream summary: The count-min sketch and its applications.'' \textit{Journal of Algorithms}, 55(1), 58-75.

\item Fan, B., Andersen, D. G., Kaminsky, M., \& Mitzenmacher, M. D. (2014). ``Cuckoo filter: Practically better than Bloom.'' In \textit{Proceedings of the 10th ACM International on Conference on Emerging Networking Experiments and Technologies} (pp. 75-88).

\item Eppstein, D., \& Goodrich, M. T. (2007). ``Space-efficient straggler identification in round-trip data streams via Newton's identities and invertible Bloom filters.'' In \textit{Algorithms and Data Structures} (pp. 637-648).

\item Tarkoma, S., Rothenberg, C. E., \& Lagerspetz, E. (2012). ``Theory and practice of Bloom filters for distributed systems.'' \textit{IEEE Communications Surveys \& Tutorials}, 14(1), 131-155.
\end{enumerate}

\section*{Systems and Implementation}

\begin{enumerate}
\setcounter{enumi}{35}
\item Bryant, R. E., \& O'Hallaron, D. R. (2015). \textit{Computer systems: A programmer's perspective} (3rd ed.). Pearson.

\item Patterson, D. A., \& Hennessy, J. L. (2017). \textit{Computer organization and design: The hardware/software interface} (5th ed.). Morgan Kaufmann.

\item Kleppmann, M. (2017). \textit{Designing data-intensive applications}. O'Reilly Media.

\item Tanenbaum, A. S., \& Van Steen, M. (2017). \textit{Distributed systems: Principles and paradigms} (3rd ed.). Pearson.
\end{enumerate}

\section*{Algorithms and Complexity}

\begin{enumerate}
\setcounter{enumi}{39}
\item Cormen, T. H., Leiserson, C. E., Rivest, R. L., \& Stein, C. (2009). \textit{Introduction to algorithms} (3rd ed.). MIT Press.

\item Sedgewick, R., \& Wayne, K. (2011). \textit{Algorithms} (4th ed.). Addison-Wesley.

\item Kleinberg, J., \& Tardos, E. (2005). \textit{Algorithm design}. Pearson.

\item Arora, S., \& Barak, B. (2009). \textit{Computational complexity: A modern approach}. Cambridge University Press.
\end{enumerate}

\section*{Additional Resources}

\begin{enumerate}
\setcounter{enumi}{43}
\item The Bernoulli-Oblivious Computing Framework Repository. Available at: \url{https://github.com/bernoulli-types}

\item IACR Cryptology ePrint Archive. Available at: \url{https://eprint.iacr.org/}

\item Proceedings of the Privacy Enhancing Technologies Symposium (PETS). Available at: \url{https://petsymposium.org/}

\item USENIX Security Symposium Proceedings. Available at:\\ \url{https://www.usenix.org/conferences/byname/108}
\end{enumerate}